---
title: "A0A7J5XRY7 Analysis"
author: "Kilicali Isildayancan"
date: "2023-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
homedir <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/chemokine_receptors/A0A7J5XRY7/"
```

# A0A7J5XRY7

## First Domain


Sequence [45-350]: TPCYQYDIYSFAQRYSPVVYILVFLVALVGNVLVLCVIRRYRNSQSGGACAFSLTDTFLLHLAISDLLLAFTLPLFAVQWAHQWVFGLVACKISGALFSLNRYSGILFLACISFDRYLAIVHAVSSGWKRSTCHAQIACALIWVGIDIYFKQVEAVKTAGNEKTLVCQVWFLENTTEWQVGLQLVSLVLGFGLPLLIMLYCYIRIFRSLCNATRRQKRKSLSLIVSLVSVFVVCWAPYNCFQLADSLHRLGLVVGGCHLGHVLDMGSLITESVGLSHCALNPLLYGFVGVKFRRELARMCKGLLGQ

### Homologies

Psi-blast search against Swiss-Prot:

```{r, echo=FALSE}
first_csvDir <- paste0(homedir,"first_psiblast.csv")
first_domain_homologs <- read.csv(first_csvDir)
colnames(first_domain_homologs) <- c("Receptor Name", "Organism", "Max Score", "Total Score", "Query Cover", "E value", "Percent Identity", "Accession Length", "Accession")

for (rowi in 1:nrow(first_domain_homologs)) {
  fullName <- first_domain_homologs[rowi,1]
  short_name <- strsplit(strsplit(first_domain_homologs$`Receptor Name`[rowi], ';')[[1]][1],"Full=")[[1]][2]
  first_domain_homologs[rowi,1] <- short_name
}

first_five <- first_domain_homologs[1:5,c(1,2,7,5,6)]
first_fifty <- first_domain_homologs[1:50,c(1,7)]
```

#### First 5 Full Homologues


```{r fiveH1, echo=FALSE}
(first_five)
```

#### First 50 Unique Homologues

```{r fiftyUnique1, echo=FALSE}
(first_fifty)
```

The similarity of the 50th match: 61% identity.

### Structure

### Motifs

```{r, echo=FALSE, fig.align='center'}
motifs_legends <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/motif_legends.png"
knitr::include_graphics(motifs_legends)
```


```{r, fig.align='center', echo=FALSE}
#ray traced, width: 11cm, height: 10cm
motifs1 <- paste0(homedir,"first_domain_motifs-colored.png")
knitr::include_graphics(motifs1)
```


### Conservation

Conservation of the protein from SwissProt with highest homology to the first domain:
CXC chemokine receptor type 3-2, UniProt ID: E9QJ73

#### Conservation in the homolog

```{r, echo=FALSE, fig.align='center'}
cons_first <- paste0(homedir,"conservation_first-domain.png")
knitr::include_graphics(cons_first)
```

#### Pairwise alignment with homolog

Looking at the conserved sites predicted by ConSurf, and cross-referencing to our first domain sequence:

```{r, echo=FALSE, fig.align='center'}
pairwise_cons <-paste0(homedir,"pairwise_homolog_first-domain.png")
knitr::include_graphics(pairwise_cons)
```

#### Binding ligands of homolog

No structures available for Chemokine receptor 3-2

#### Superimposition

RMSD: 1.482 \AA

```{r, fig.align='center', echo=FALSE}
#ray traced, width: 11cm, height: 10cm
motifs1 <- paste0(homedir,"consurf_superimposed_first-domain.png")
knitr::include_graphics(motifs1)
```

\* Both are AlphaFold predicted structures

## Second domain

Sequence [407-687]: YEYKDEIEARVSKAVLLPILYSLVLVVGLLGNVFLLAVLVQKRFRSWSISDIFILHLSVSDLLLLVSLPLWAAEAAQISGWCCGGFLCKINCYLSIVHSTQRFSQKNPRSAHICCLLVWLLSVIFSIPDWIFLSVKTDAGKRLCNSYAQSISTSQLVSRLLHLTLGFLLPSLALIICSSFILLRMKRGSECPQKKRAGRLLLLLVVVFFLCWMPYHIILFVDTIRKSSKERDGGLSRNLKSSQITALMFTSAVGCLHACLRPLLYFGLCGNFRERTLAMLT

### Homologies

Psi-blast search against Swiss-Prot:

```{r, echo=FALSE}
second_csvDir <-paste0(homedir, "second_psiblast.csv")
second_domain_homologs <- read.csv(second_csvDir)
colnames(second_domain_homologs) <- c("Receptor Name", "Organism", "Max Score", "Total Score", "Query Cover", "E value", "Percent Identity", "Accession Length", "Accession")

for (rowi in 1:nrow(second_domain_homologs)) {
  fullName <- second_domain_homologs[rowi,1]
  short_name <- strsplit(strsplit(second_domain_homologs$`Receptor Name`[rowi], ';')[[1]][1],"Full=")[[1]][2]
  second_domain_homologs[rowi,1] <- short_name
}

first_five <- second_domain_homologs[1:5,c(1,2,7,5,6)]
first_fifty <- second_domain_homologs[1:50,c(1,7)]
```


#### First 5 Full Homologues


```{r fiveH2, echo=FALSE}
(first_five)
```



#### First 50 Unique Homologues

```{r fiftyUnique2, echo=FALSE}
(first_fifty)
```

The similarity of the 50th match: 61% identity.

### Structure

### Motifs

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics(motifs_legends)
```

```{r, fig.align='center', echo=FALSE}
#ray traced, width: 11cm, height: 10cm
motifs2 <- paste0(homedir,"second_domain_motifs-colored.png")
knitr::include_graphics(motifs2)
```

### Conservation

Conservation of the protein from SwissProt with highest homology to the second domain:
C-C chemokine receptor type 5, UniProt ID: Q2HJ17

#### Conservation with homolog

```{r, echo=FALSE, fig.align='center'}
cons_first <- paste0(homedir,"conservation_second-domain.png")
knitr::include_graphics(cons_first)
```

#### Pairwise alignment with homolog

```{r, echo=FALSE, fig.align='center'}
pairwise_cons <- paste0(homedir,"pairwise_homolog_second-domain.png")
knitr::include_graphics(pairwise_cons)
```


#### Binding ligands of homolog

5UIW is with natural ligand (CCR5),
6AKX, 6AKY is a non-natural ligand (Compound 21, 34),

#### Ligand superimposition of domain and homolog with ligand

RMSD with the ligand-bound receptor: 0.506 \AA

```{r, echo=FALSE, fig.align='center'}
ligand_superimp2 <-paste0(homedir,"second_domain_ligand-superimposed.png")
knitr::include_graphics(ligand_superimp2)
```


#### Superimposition

RMSD: 1.733 \AA

```{r, fig.align='center', echo=FALSE}
consurf_superimpo2 <- paste0(homedir,"consurf_superimposed_second-domain.png")
knitr::include_graphics(consurf_superimpo2)
```

## Relative domains

### Pairwise alignment

```{r, fig.align='center', echo=FALSE}
pwAli <- paste0(homedir,"pairwise_alignment_needle.png")
knitr::include_graphics(pwAli)
```

## Taxonomy Cluster

```{r, fig.align='center', echo=FALSE}
taxC <- paste0(homedir,"cluster149.png")
knitr::include_graphics(taxC)
```

## Genomic sequence

Does not exist in a genome database viewer.

