---
title: "Allatostatin-C-like Receptor Analysis"
author: "Kilicali Isildayancan"
date: "2023-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# A0A891XI80

## First Domain


Sequence [52-342]: PTLLLITQVLCAAVCIVGLLGNSLVIYVVLRFSKMQTVTNMYIVNLAVADECFLIGVPF
LLATMNLRIWPFGSIMCKAYMTTTSINQFTSSIFLTIMSADRYIAVCHPISAPKVRTPLISKVVSASAWTASALLMIP
IFIYANIMTNENVVSCNILWPESENVSSQGAFTLYTFILGFSIPLMLISVFYILVIRKLKTVGPKNKSKEKKKSHRKV
TKLVLTVITVYVLCWLPYWLTQMALIFTPPKQCQSRVSITVSLLAGCLSYSNSAMNPILYAFLSDNFKKSFLKACT

### Homologies

Psi-blast search against Swiss-Prot:


#### First 5 Full Homologues


```{r fiveH1, echo=FALSE}
receptorNames <- c("Somatostatin receptor type 2", "Somatostatin receptor type 1", "Somatostatin receptor type 1", "Somatostatin receptor type 1", "Somatostatin receptor type 1")
organisms <- c("Homo sapiens","Rattus norvegicus","Mus musculus","Homo sapiens","Canis lupus familiaris")
identities <- c("47%","43%","43%","43%","43%")
positives <- c("68%","68%","68%","68%","68%")
fiveH <- data.frame(receptorNames, identities, positives, organisms)
(fiveH)
```

#### First 50 Unique Homologues

```{r fiftyUnique1, echo=FALSE}
uniqReceptorNames <- c("Somatostatin receptor type 2","Somatostatin receptor type 1","Somatostatin receptor type 4","Somatostatin receptor type 5","Somatostatin receptor type 3","Nociceptin receptor", "Kappa-type opioid receptor", "Neuropeptides B/W receptor type 2", "Delta-type opioid receptor", "Mu-type opioid receptor", "Neuropeptides B/W receptor type 1", "Galanin receptor")
(uniqReceptorNames)
```

The similarity of the 50th match: 33% identity, 57% positives

### Structure

### Motifs

```{r, echo=FALSE, fig.align='center'}
motifs_legends <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/motif_legends.png"
knitr::include_graphics(motifs_legends)
```


```{r, fig.align='center', echo=FALSE}
#ray traced, width: 11cm, height: 10cm
motifs1 <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/allatostatin-C-like/first_domain_motifs-colored.png"
knitr::include_graphics(motifs1)
```


### Conservation

Conservation of the protein from SwissProt with highest homology to the first domain:
Somatostatin receptor type 2, UniProt ID: P30874

#### Conservation in the homolog

```{r, echo=FALSE, fig.align='center'}
cons_first <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/allatostatin-C-like/conservation_first-domain.png"
knitr::include_graphics(cons_first)
```

#### Pairwise alignment with homolog

Looking at the conserved sites predicted by ConSurf, and cross-referencing to our first domain sequence:

```{r, echo=FALSE, fig.align='center'}
pairwise_cons <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/allatostatin-C-like/pairwise_homolog_first-domain.png"
knitr::include_graphics(pairwise_cons)
```

#### Binding ligands of homolog

7T10 is with natural ligand,
7WIC is a non-natural ligand

#### Ligand superimposition of domain and homolog with ligand


#### Superimposition

RMSD: 0.895

```{r, fig.align='center', echo=FALSE}
#ray traced, width: 11cm, height: 10cm
motifs1 <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/allatostatin-C-like/consurf_superimposed_first_domain.png"
knitr::include_graphics(motifs1)
```

\* Both are AlphaFold predicted structures

## Second domain

Sequence [448-700]: TNLLIINLALADLLFIVFCVPFTATDYMLPYWPFGDVWCKIVQYLIVVTAYASVYTLV
LMSLDRFLAVVHPIASISVRTERNAVTAIGVTWVVILLSSVPVYLSHGVTNYIYLSSEHTACVFLDDYNKPVFQILFF
ATSYVIPLALICGLYLCMLMRLWRGVAPGGHVSAESRRGRKRVTRMVVVVVAIFAVCWFPIQMILVLKSVDRYEITNT
SVMVQIVSHVLAYMNSCVNPILYAFLSENFRKAFRKVI

### Homologies

Psi-blast search against Swiss-Prot:


#### First 5 Full Homologues


```{r fiveH2, echo=FALSE}
receptorNames <- c("Allatostatin-A receptor", "Galanin receptor 2a", "Galanin receptor type 2", "Galanin receptor 2b", "Somatostatin receptor type 1")
organisms <- c("Bombyx mori","Danio rerio","Mus musculus","Danio rerio","Rattus norvegicus")
identities <- c("70%","40%","40%","39%","40%")
positives <- c("85%","63%","61%","62%","61%")
fiveH <- data.frame(receptorNames, identities, positives, organisms)
(fiveH)
```



#### First 50 Unique Homologues

```{r fiftyUnique2, echo=FALSE}
uniqReceptorNames <- c("Allatostatin-A receptor", "Galanin receptor 2a", "Galanin receptor type 2", "Galanin receptor 2b", "Somatostatin receptor type 1", "Galanin receptor type 3", "Galanin receptor type 1", "Nociceptin receptor", "Somatostatin receptor type 5", "Mu-type opioid receptor", "KiSS-1 receptor", "Kappa-type opioid receptor", "Somatostatin receptor type 2", "G-protein coupled receptor 54 (orphan)", "Somatostatin receptor type 3", "Somatostatin receptor type 4","Delta-type opioid receptor")
(uniqReceptorNames)
```

The similarity of the 50th match: 35% identity, 53% positives

### Structure

### Motifs

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics(motifs_legends)
```

```{r, fig.align='center', echo=FALSE}
#ray traced, width: 19cm, height: 15cm
motifs2 <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/allatostatin-C-like/second_domain_motifs-colored.png"
knitr::include_graphics(motifs2)
```

### Conservation

Conservation of the protein from SwissProt with highest homology to the second domain:
Allatostatin-A receptor, UniProt ID: Q8WPA2

#### Conservation with homolog

#### Pairwise alignment with homolog

```{r, echo=FALSE, fig.align='center'}
pairwise_cons <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/allatostatin-C-like/pairwise_homolog_second-domain.png"
knitr::include_graphics(pairwise_cons)
```

#### Superimposition

RMSD: 0.895

```{r, fig.align='center', echo=FALSE}
consurf_superimpo2 <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/allatostatin-C-like/consurf_superimposed_second_domain.png"
knitr::include_graphics(consurf_superimpo2)
```

## Relative domains

### Pairwise alignment

```{r, fig.align='center', echo=FALSE}
pwAli <- "/home/kilicali/multi-domain_gpcr/mGPCR_analysis/allatostatin-C-like/pairwise_alignment_needle.png"
knitr::include_graphics(pwAli)
```

## Genomic sequence

Does not exist in a genome database viewer.

