---
title: "Class A Receptor Taxonomical Analysis"
author: "Kilicali Isildayancan"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```




# Taxonomical Analysis


```{r startup, echo=FALSE, warning=FALSE, message=FALSE}
library('ggtree')
library('ggplot2')
library('ggimage')
library('RColorBrewer')
library('jsonlite')
json_uniprot_directory <- '/home/kilicali/multi-domain_gpcr/datafiles/ur100_chemokines_7TM.json'
json_cluster_directory <- '/home/kilicali/multi-domain_gpcr/datafiles/clusters/chemokine_receptors/7TM/fasta_clustered/five_plus_clusters.json'
jsondata <- jsonlite::fromJSON(json_uniprot_directory)
json_clusters <- jsonlite::fromJSON(json_cluster_directory)
```


```{r}
cluster <- json_clusters[[8]]
taxIds <-unique(jsondata$organism$taxonId)
in_cluster_taxIds <-unique(jsondata$organism$taxonId[jsondata$primaryAccession %in% cluster])
taxDB <- taxize::classification(taxIds, db = 'ncbi')
cluster_taxDB <- taxize::classification(in_cluster_taxIds, db = 'ncbi')
taxClass <- taxize::classification(unique(taxIds), db = 'ncbi')
fullTree <- taxize::class2tree(taxDB, check=FALSE)

ggtree::ggtree(fullTree$phylo)+ggtree::geom_nodelab(nudge_x=-5.5,nudge_y=0.35)+ggtree::geom_tiplab()+ggtree::xlim(-10,90)
```

```{r}
clusterTree <- taxize::class2tree(cluster_taxDB, check=FALSE)
ggtree::ggtree(clusterTree$phylo)+ggtree::geom_nodelab(nudge_x=-5.5,nudge_y=0.35)+ggtree::geom_tiplab()+ggtree::xlim(-10,90)
```

### Kingdoms

-   There are `r length(unique(allKs))` kingdoms, namely: `r unique(allKs)`.

### Phylum

-   There are `r length (unique(allPs))` phylums, namely: `r unique(allPs)`.

Here is a visual representation of all the phylums present (which covers all the lower-level taxa):

```{r treeP, fig.align='center', fig.width=15}
ggtree::ggtree(phylumTree$phylo)+ggtree::geom_nodelab(nudge_x=-5.5,nudge_y=0.35)+ggtree::geom_tiplab()+ggtree::xlim(-10,90)
```

Also the histogram distribution:

```{r histPhylum, fig.align='center', fig.width=15}
# Create the bar plot
ggplot2::ggplot(phylumDF, aes(x = allPs, y = Freq, fill = allPs)) + 
  geom_bar(stat = "identity") +
  labs(title = "diChemokines in Organisms by Phylum", x = "Phylum", y = "Frequency") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=12))
```

The chordates represent `r sum(allPs=='Chordata')/length(allPs)*100` of all the phylums in the reference proteomes with a total of `r sum(allPs=='Chordata')`.

## Mid-level taxonomy (Order)

We are considering only order here, and looking for primates in particular.


```{r orderTable}
(table(allOs))
```


## Low-level taxonomy (Family and Species)


```{r orderTable}
(table(allFs))
```



### Families

```{r family_tree, fig.align='center'}

ggtree::ggtree(famTree$phylo)+ggtree::geom_nodelab(nudge_x=-5.5,nudge_y=0.35)+ggtree::geom_tiplab()+ggtree::xlim(-10,90)
```

## Absent children

```{r}
order_id <- "9443"
taxize::downstream(order_id, db='ncbi', downto='species')
```


### Orders of model organisms

Considered model organisms (and respective orders): if they are in our protein phylogeny

-   Mus musculus (Rodentia): `r ("Rodentia" %in% allOs)`

-   Rattus norvegicus (Rodentia): `r ("Rodentia" %in% allOs)`

-   Drosophila melanogaster (Drosophilidae): `r ("Drosophilidae" %in% allOs)`

-   Bos taurus (Artiodactyla): `r ("Artiodactyla" %in% allOs)`

-   Callosobruchus maculatus (Coleoptera): `r ("Coleoptera" %in% allOs)`

-   Danio rerio (Cypriniformes): `r ("Cypriniformes" %in% allOs)`

-   Fundulus heteroclitus (Cyprinodontiformes) \~not a "model" : `r ("Cyprinodontiformes" %in% allOs)`

-   Nothobranchius furzeri (Cyprinodontiformes): `r ("Cyprinodontiformes" %in% allOs)`

-   Oryzias latipes (Beloniformes): `r ("Beloniformes" %in% allOs)`

-   Anolis carolinensis (Squamata): `r ("Squamata" %in% allOs)`

-   Gallus gallus (Galliformes): `r ("Galliformes" %in% allOs)` (UPI001F01E243 is in Gallus gallus?)

-   Taeniopygia castanotis (Passeriformes): `r ("Passeriformes" %in% allOs)`

-   Xenopus laevis (Anura): `r ("Anura" %in% allOs)`

-   Xenopus tropicalis (Anura): `r ("Anura" %in% allOs)`

-   Saccharomyces cerevisiae (Saccharomycetales): `r ("Saccharomycetales" %in% allOs)`

-   Any primate might be good to start with (Primates): `r ("Primates" %in% allOs)`

-   Caenorhabditis elegans (Rhabditida): `r ("Rhabditida" %in% allOs)`

-   Strongylocentrotus purpuratus (Echinoida): `r ("Echinoida" %in% allOs)`




