---
title: "Reference Proteomes Taxonomical Analysis"
author: "Kilicali Isildayancan"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{=html}
<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Taxonomical Analysis

## First look

```{r startup, echo=FALSE, warning=FALSE, message=FALSE}
library('ggtree')
library('ggplot2')
library('ggimage')
library('RColorBrewer')
direcTax <- '/home/kilicali/multi-domain_gpcr/phylo/tax-refprot'
fileTax <- read.csv(direcTax, sep = '\t')
taxIds <- as.numeric(fileTax[[1]])
```

```{r fetchingInfo, warning=FALSE}
taxDB <- taxize::classification(taxIds, db = 'ncbi')
```

## High-level taxonomy (Superkingdoms, kingdoms, and phylum)

```{r low_level3s}
allSKs<-c()
allKs <- c()
allPs <- c()
idPs <- c()
for (taxo in taxDB) {
  superkingdom<-taxo$name[taxo$rank=='superkingdom']
  allSKs <- c(allSKs, superkingdom)
  kingdom<-taxo$name[taxo$rank=='kingdom']
  allKs <- c(allKs, kingdom)
  phylum<-taxo$name[taxo$rank=='phylum']
  allPs <- c(allPs, phylum)
  tmpid <- taxo$id[taxo$rank=='phylum']
  idPs <- unique(c(idPs, tmpid))
}

idPs <- as.numeric(idPs)
```

### Superkingdoms

-   There are `r length(unique(allSKs))` superkingdom(s), namely: `r unique(allSKs)`. Eukaryotes represent `r sum(allSKs=='Eukaryota')/length(allSKs)*100` (There are `r sum(allSKs=='Bacteria')` from Bacteria).

### Kingdoms

-   There are `r length(unique(allKs))` kingdoms, namely: `r unique(allKs)`. Here is the distribution:

```{r histKingdoms, fig.align='center', fig.width=15}
kingdomDF <- as.data.frame(table(allKs))
# Create the bar plot
ggplot2::ggplot(kingdomDF, aes(x = allKs, y = Freq, fill = allKs)) + 
  geom_bar(stat = "identity") +
  labs(title = "Categorical Frequencies", x = "Kingdom", y = "Frequency") +
  theme(axis.text.x=element_blank())
```

### Phylum

-   There are `r length (unique(allPs))` phylums, namely: `r unique(allPs)`.

Here is a visual representation of all the phylums present (which covers all the lower-level taxa):

```{r treeP, fig.align='center', fig.width=10, fig.height=12}
phylumClass<-taxize::classification(idPs,'ncbi')
phylumTree <- taxize::class2tree(phylumClass)
ggtree::ggtree(phylumTree$phylo)+ggtree::geom_nodelab(nudge_x=-5.5,nudge_y=0.35)+ggtree::geom_tiplab()+ggtree::xlim(-10,90)
```

Also the histogram distribution:

```{r histPhylum, fig.align='center', fig.width=20}

phylumDF <- as.data.frame(table(allPs))
# Create the bar plot
ggplot2::ggplot(phylumDF, aes(x = allPs, y = Freq, fill = allPs)) + 
  geom_bar(stat = "identity") +
  labs(title = "diChemokines in Organisms by Phylum", x = "Phylum", y = "Frequency") +
  geom_text(aes(label = allPs),fontface = "bold",nudge_y=4.2,size=5) +
  theme(axis.text.x=element_blank())
```

The chordates represent `r sum(allPs=='Chordata')/length(allPs)*100` of all the phylums in the reference proteomes with a total of `r sum(allPs=='Chordata')`.

## Mid-level taxonomy (Order)

We are considering only order here, and looking for the phylum of Chordata in particular.

```{r mid_level}
allOs <- c()
phylumDB <- data.frame()
for (taxo in taxDB) {
  order<-taxo$name[taxo$rank=='order']
  allOs <- c(allOs, order)
}
```

```{r histOrder}
orderDF <- as.data.frame(table(allOs))
split1ord <- data.frame(allOs=orderDF$allOs[orderDF$Freq>30],Freq=orderDF$Freq[orderDF$Freq>30])
split2ord <- data.frame(allOs=orderDF$allOs[orderDF$Freq<=30 & orderDF$Freq>=10],Freq=orderDF$Freq[orderDF$Freq<=30 & orderDF$Freq>=10])
split3ord <- data.frame(allOs=orderDF$allOs[orderDF$Freq<=10 & orderDF$Freq>=5],Freq=orderDF$Freq[orderDF$Freq<=10 & orderDF$Freq>=5])
split4ord <- data.frame(allOs=orderDF$allOs[orderDF$Freq<5],Freq=orderDF$Freq[orderDF$Freq<5])
```

```{r ordGT30, fig.align='center', fig.width=12}
# Create the bar plot
ggplot2::ggplot(split1ord, aes(x = allOs, y = Freq, fill = allOs)) + 
  geom_bar(stat = "identity") +
  labs(title = "diChemokines in Organisms by Order", x = "Order", y = "Frequency") +
  theme_minimal(base_size = 15)+
  theme(axis.text.x=element_text(size=12,face='bold'))
```

```{r ordBTW10and30, fig.align='center', fig.width=20}
# Create the bar plot
ggplot2::ggplot(split2ord, aes(x = allOs, y = Freq, fill = allOs)) + 
  geom_bar(stat = "identity") +
  labs(title = "diChemokines in Organisms by Order", x = "Order", y = "Frequency") +
  geom_text(aes(label=allOs),nudge_y=.3, size=6)+
  theme_minimal() +
  theme(axis.text.x=element_blank())
```

```{r ordsplit3, fig.align='center', fig.width=25}
# Create the bar plot
ggplot2::ggplot(split3ord, aes(x = allOs, y = Freq, fill = allOs)) + 
  geom_bar(stat = "identity") +
  labs(title = "diChemokines in Organisms by Order", x = "Order", y = "Frequency") +
  geom_text(aes(label=allOs),nudge_y=.1, size=6, check_overlap=TRUE)+
  theme_minimal() +
  theme(axis.text.x=element_blank())
```

## Low-level taxonomy (Family and Species)

```{r lo_level}
allFams <- c()
idFs <- c()
for (taxo in taxDB) {
  family<-taxo$name[taxo$rank=='family']
  allFams <- c(allFams, family)
  if ('Cricetidae' %in% taxo$name) {
    tmpid <- taxo$id[taxo$rank=="species"]
    idFs <- unique(c(idFs, tmpid))
  }
  
}

idFs <- as.numeric(idFs)
```

```{r histFamily}
familyDF <- as.data.frame(table(allFams))
split1f <- data.frame(allFams=familyDF$allFams[familyDF$Freq>30],Freq=familyDF$Freq[familyDF$Freq>30])
split2f <- data.frame(allFams=familyDF$allFams[familyDF$Freq<30 & familyDF$Freq>10],Freq=familyDF$Freq[familyDF$Freq<30 & familyDF$Freq>10])
split3f <- data.frame(allFams=familyDF$allFams[familyDF$Freq<10 & familyDF$Freq>5],Freq=familyDF$Freq[familyDF$Freq<10 & familyDF$Freq>5])
split4f <- data.frame(allFams=familyDF$allFams[familyDF$Freq<5],Freq=familyDF$Freq[familyDF$Freq<5])
```

```{r famGT30, fig.align='center', fig.width=10}
# Create the bar plot
ggplot2::ggplot(split1f, aes(x = allFams, y = Freq, fill = allFams)) + 
  geom_bar(stat = "identity") +
  labs(title = "diChemokines in Organisms by Family", x = "Family", y = "Frequency") +
  theme_minimal()
```

```{r famBTW3and5, fig.align='center', fig.width=20}
# Create the bar plot
ggplot2::ggplot(split2f, aes(x = allFams, y = Freq, fill = allFams)) + 
  geom_bar(stat = "identity") +
  labs(title = "diChemokines in Organisms by Family", x = "Family", y = "Frequency") +
  geom_text(aes(label=allFams),nudge_y=.3, size=6,check_overlap=TRUE)+
  theme_minimal(base_size = 25) +
  theme(axis.text.x=element_blank())
```

```{r famLT3}
#There is a need to split further this group if need be, but I am ignoring it for the current analysis
# Create the bar plot
```

### Cricetidae

```{r Cricetidae, fig.align='center', fig.width=10}
famClass<-taxize::classification(idFs,'ncbi')
famTree <- taxize::class2tree(famClass,check=FALSE)
ggtree::ggtree(famTree$phylo)+ggtree::geom_nodelab(nudge_x=-4,nudge_y=0.065)+ggtree::geom_tiplab()+ggtree::xlim(-10,90)
```

### Families of model organisms

Considered model organisms (and respective orders): if they are in our protein phylogeny

-   Mus musculus (Rodentia): `r ("Rodentia" %in% allOs)`

-   Rattus norvegicus (Rodentia): `r ("Rodentia" %in% allOs)`

-   Drosophila melanogaster (Drosophilidae): `r ("Drosophilidae" %in% allOs)`

-   Bos taurus (Artiodactyla): `r ("Artiodactyla" %in% allOs)`

-   Callosobruchus maculatus (Coleoptera): `r ("Coleoptera" %in% allOs)`

-   Danio rerio (Cypriniformes): `r ("Cypriniformes" %in% allOs)`

-   Fundulus heteroclitus (Cyprinodontiformes) \~not a "model" : `r ("Cyprinodontiformes" %in% allOs)`

-   Nothobranchius furzeri (Cyprinodontiformes): `r ("Cyprinodontiformes" %in% allOs)`

-   Oryzias latipes (Beloniformes): `r ("Beloniformes" %in% allOs)`

-   Anolis carolinensis (Squamata): `r ("Squamata" %in% allOs)`

-   Gallus gallus (Galliformes): `r ("Galliformes" %in% allOs)` (UPI001F01E243 is in Gallus gallus?)

-   Taeniopygia castanotis (Passeriformes): `r ("Passeriformes" %in% allOs)`

-   Xenopus laevis (Anura): `r ("Anura" %in% allOs)`

-   Xenopus tropicalis (Anura): `r ("Anura" %in% allOs)`

-   Saccharomyces cerevisiae (Saccharomycetales): `r ("Saccharomycetales" %in% allOs)`

-   Any primate might be good to start with (Primates): `r ("Primates" %in% allOs)`

-   Caenorhabditis elegans (Rhabditida): `r ("Rhabditida" %in% allOs)`

-   Strongylocentrotus purpuratus (Echinoida): `r ("Echinoida" %in% allOs)`

There are low amount of hits in Muridae, but it can be interesting since it has Rattus and Mus in it as geni (mice and rats as model organisms)
