---
title: "Chemokine Receptor Taxonomical Analysis"
author: "Kilicali Isildayancan"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Taxonomical Analysis

## First look

```{r startup, echo=FALSE, warning=FALSE, message=FALSE}
library('ggtree')
library('ggplot2')
library('ggimage')
library('RColorBrewer')
direcProts <- '/home/kilicali/multi-domain_gpcr/focus_chemokines/chemokines_preliminary.tsv'
direcTax <- '/home/kilicali/multi-domain_gpcr/focus_chemokines/tax-chemo-new'
direcImage <- '/home/kilicali/multi-domain_gpcr/focus_chemokines/trees'
```

First we get the TSV file from the HMMSEARCH output parser. This file contains names of proteins (e.g. UniRef90_A0A1B0CRB2), amount of chemokine receptor domain counts, and lastly the taxonomic identifier (e.g. 9606 is for "Homo Sapiens")

```{r directories, warning=FALSE}
fileProts <- read.csv(direcProts, sep = '\t')
fileTax <- read.csv(direcTax, sep = '\t')

#Protein ID & domain count
protNames <- fileProts$Name
domainCount <- fileProts$HitCount

#Protein ID & tax id
taxProts <- fileTax$ID
taxIds <- as.numeric(fileTax$Tax)
```

```{r preFiltering}
#This is done because current TaxIDs do not span all of the prot names, remove when doing it with proper tax file
filt_protNames <- protNames[protNames %in% taxProts]
domainCount <- domainCount[protNames %in% filt_protNames]
protNames <- filt_protNames
```

### Domains

Here, we can see that we have `r length(unique(protNames))` unique proteins with a minimum of `r min(domainCount)` and a maximum of `r max(domainCount)` domain hits. The distribution of the number of domain frequencies are as follows:

```{r histoDomainCount, echo=FALSE, fig.align='center', fig.width=10}
ddist <- hist(domainCount,
     breaks=c(1,2,3,4,5,6,7,8,9,10,11,12,13),
     xlim=range(1,13),
     col=brewer.pal(n=11,'RdBu'),
     xlab="Number of domains",
     main='Distribution chemokine receptor domain counts',xaxt="n")
axis(side=1,at=ddist$mids,labels=seq(2,13))
```
